### 2021-04-26 manuscript.tex
Initial Overleaf Import

%%
%% This is file `./samples/longsample.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% apa7.dtx  (with options: `longsample')
%% ----------------------------------------------------------------------
%% 
%% apa7 - A LaTeX class for formatting documents in compliance with the
%% American Psychological Association's Publication Manual, 7th edition
%% 
%% Copyright (C) 2019 by Daniel A. Weiss <daniel.weiss.led at gmail.com>
%% 
%% This work may be distributed and/or modified under the
%% conditions of the LaTeX Project Public License (LPPL), either
%% version 1.3c of this license or (at your option) any later
%% version.  The latest version of this license is in the file:
%% 
%% http://www.latex-project.org/lppl.txt
%% 
%% Users may freely modify these files without permission, as long as the
%% copyright line and this statement are maintained intact.
%% 
%% This work is not endorsed by, affiliated with, or probably even known
%% by, the American Psychological Association.
%% 
%% ----------------------------------------------------------------------
%% 
\documentclass[man]{apa7}

\usepackage[american]{babel}
\usepackage{nicefrac}
\usepackage{todonotes}
\usepackage{epigraph}
% For sorting references from same year in order instead of on title
\newcommand{\SR}[1]{} %for EJ's bibtex file
\newcommand{\EJ}[1]{\todo[inline, color=green]{  #1 }}
\newcommand{\JR}[1]{\todo[inline, color=gray]{  #1 }}
\newcommand{\MK}[1]{\todo[inline, color=red]{  #1 }}

\usepackage{graphicx}
\graphicspath{ {images/} }

\usepackage{csquotes}
\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber]{biblatex}
\DeclareLanguageMapping{american}{american-apa}
\addbibresource{bibliography.bib}

\title{Trust Me, I'm a Bayesian: Building Back Trust in Science Using Bayes}
\shorttitle{Bayes in Science Education}

\author{Joshua M. Rosenberg}
\affiliation{University of Tennessee, Knoxville}

\author{Marcus Kubsch}
\affiliation{IPN - Leibniz Institute for Science and Mathematics Education}

\author{Eric-Jan Wagenmakers}
\affiliation{University of Amsterdam}

\leftheader{Weiss}

\abstract{

Uncertainty is ubiquitous in science and this uncertainty takes many forms, but the rhetoric around scientific knowledge and negotiating standards of evidence is often portrayed as immutable. Adding to this problem around the limited recognition of the tentative nature of science, when scientists try to communicate their uncertainty, research has shown us that many—including both learners and experts—struggle to interpret the statistics that are often used to make inferences from scientific data. We argue that a Bayesian approach can offer a different foundation and language around a core but sometimes unsettling tenet and persistent feature, uncertainty. Central to this conceptualization is viewing uncertainty probabilistically, as probabilities can provide a language for expressing the degree of uncertainty in our knowledge in a wide range of domains. In this essay, we a) briefly introduce Bayes theorem and wider ideas around Bayesian epistemology, b) elaborate on the importance of probability and uncertainty in science and science education and review how people learn about these concepts and where they face difficulties, and c) provide conceptual and practical examples within the JASP statistical software of how Bayes could help to address these issues. We conclude with implications and directions for future research focused around bolstering trust in science outside of classroom settings (for which Bayesian approaches could serve as a crosscutting part of science literacy) and in pre-collegiate science classrooms, where Bayes can provide a coherent approach to engaging in numerous science practices such as arguing from evidence and analyzing and interpreting data in a principled and flexible way.
}

\keywords{APA style, demonstration}

\authornote{
   \addORCIDlink{Joshua M. Rosenberg}{0000-0003-2170-0447}
   \addORCIDlink{Marcus Kubsch}{0000-0000-0000-0000}
   \addORCIDlink{Eric-Jan Wagenmakers}{0000-0003-1596-1034} 
}

\begin{document}
\maketitle

\setlength{\epigraphwidth}{1.0\textwidth} % set for individual epigraph
\epigraph{It is seen in this essay that the theory of probabilities is at bottom only common sense reduced to calculus; it makes us appreciate with exactitute that which exact minds feel by a sort of instinct without being able ofttimes to give a reason for it . . . we shall see that there is no science more worthy of our meditations, and that no more useful one could be incorporated in the system of public instruction.}{Pierre-Simon Laplace, 1829}
\setlength{\epigraphwidth}{0.4\textwidth} % return to default for the next epigraphs
%\nocite{Laplace18291902} %[p. 196]

\EJ{This is a comment.}

Uncertainty is ubiquitous in science and this uncertainty takes many forms. For example, measurement uncertainty is present when there are differences (small or large) in the repeated measures of the same object or phenomenon \parencite{pls03} or when the object that is measured, like a plant, is inherently variable in its characteristics \parencite{pls03}. Another form of uncertainty is its role within explanatory theories, such as those associated with evolution and quantum physics, for which central models take the form of probablistic statments, i.e., uncertainty is inherent in these models (Gigerenzer, 2000).

Negotiating these many forms of uncertainty in argumentation—and building toward consensus at the same time that uncertainty is present—is a key part of the scientific process (Thagard, 2000). In consequence, some philosophers of science have suggested that science is a process that builds increasingly better models which allows us to make increasingly more accurate predictions about the world (Giere, 2010; Godfrey-Smith, 2006; Nersessian, 2002; Reichenbach, 1977).

However, while scientists and philosophers of science may be aware of the uncertainty (and limitations) surrounding scientific knowledge and negotiating standards of evidence interpretation, scientific knowledge is still often portrayed as immutable in educational contexts (Duschl et al., 1990). In consequence, trust in science can erode when scientists individually or collectively change their views, such as during the first phase of the COVID-19 crisis in Germany, when scientists faced serious criticism from politicians and the media for changing their opinion based on new information. A similar tension was present in the United States and around the world regarding potential COVID-19 medication, when initial scientific data and changing interpretations of its utility collided with citizens’ skepticism (and, partisan political views).

Adding to this problem around the limited recognition of the tentative nature of science, when scientists try to communicate their uncertainty, research has shown us that many—including learners and experts alike—struggle to interpret statistics such as confidence intervals or error bars (Gigerenzer et al. 2004; Sedlmeier, 2007; Tversky & Kahneman, 1977). Similarly, difficulties in understanding the language of probability can contribute to a struggle on the part of students when learning about concepts such as evolution of nuclear decay that are modelled in terms of probabilistic statements (e.g., Fiedler et al., 2019).

Based on accumulating evidence from cognitive psychology, neuroscience, and education, we build on and extend beyond the science and engineering practice of argumentation the foundational work of science education (e.g., Nussbaum, 2011; Szu & Osborne, 2012) and statistics education scholars (e.g., Albert, 2002; Bolstad, 2002; Hoegh et al., 2020; Hu, 2020; Martignon and Erickson, 2014, Sedlmeier, 2007; Witmer, 2017) to argue for a relatively novel approach to uncertainty and probability. Specifically, we argue that it is helpful to consider uncertainty in terms of a Bayesian epistemological framework to help people to understand uncertainty and probability better and to engage with them in a way that leverages their existing ways of thinking. Central to this conceptualization is viewing uncertainty probabilistically, as probabilities can provide a language for expressing the degree of uncertainty in our knowledge in a wide range of domains (Gopnik, 2012). In this way, we make this conjecture not only on the merits of Bayes Theorem as a mathematical or statistical tool, but based on a body of psychological and educational research that shows that such an approach is an intuitive way for people to think about data and evidence in the world (Gopnik, 2012) and in classroom contexts (Leuders & Loibl, 2020; Nussbaum, 2011; Sedlmeier, 2007; Szu & Osborne, 2012; Witmer, 2017).
The remainder of this paper is structured in the following way. We will first briefly introduce Bayes theorem and wider ideas around Bayesian epistemology. Next, we will elaborate on the importance of probability and uncertainty in science (education) and review how people learn about these concepts and where they face difficulties. Then, we will provide examples of how Bayes could help to address these issues before concluding with implications and directions for future research.

\section{Primer on Bayes Theorem and Bayesian Views of Epistemology}

In essence, Bayes’ theorem provides a framework for updating one’s beliefs about the world from observations in a way that takes prior knowledge into account in a principled manner. Bayes theorem (Equation 1) states that our knowledge about the world after making an observation can be expressed as a probability distribution over the possible outcomes of the phenomenon we observed—the posterior distribution. To calculate the posterior, we express our knowledge about the world before the observation in terms of a probability distribution, the prior, and multiply that with the likelihood which summarizes the information in our observation. These are then divided by the average likelihood to yield a posterior distribution that sums to one.:

Posterior = Prior x LikelihoodAverage Likelihood

\EJ{So I thought about introducing this more elaborately with an example where interest centers on what proportion of students is using
Windows vs. the Mac. Data from the first 10 students are $\{W, W, M, W, M, W, W, M, W, M\}$. I'm using a beta$(2,2)$ prior. The results are shown in the two figures below.}

These are big ideas, and, fortunately, there are a number of intuitive explanations of Bayesian methods, including Kurt (2019), Kruschke and Liddell (2018a) Wagenmakers et al. (2016). A recent example from physics can serve to illustrate this principle in the context of science education. In 2011, physicists reported that they had observed elementary particles moving faster than the speed of light (Brumfiel, 2011) - and thus defying Einstein’s theory of relativity. However, instead of celebrating a historic discovery, the scientists started to look for what had gone wrong in their observations.

How can Bayes’ theorem help to make sense of this? With nearly a century of accumulated evidence that the theory of relativity holds, scientists had a very strong prior belief about the speed of light. To change this belief, extraordinary evidence would have been required. Thus, a single instance of apparently faster-than light elementary particles does not give a likelihood strong enough to change the prior in a meaningful way and in consequence the posterior, the knowledge about the world after making an observation, does not change in a meaningful way.

Thus, rather than changing their belief about the speed of light, scientists looked for things that could have gone wrong with the measurement. This example shows how a qualitative take on Bayes’ theorem can explain and predict many aspects of how animals (Okasha et al., 2012), people (Griffiths et al. 2008, Tenenbaum et al. 2006), and organisations (Kruschke et al., 2012) learn from observations. Recently, advances in computational power and techniques have started the more and more widespread application of this principle in a quantitative way that allows researchers to precisely express their knowledge about the world and then learn from observations.

Expressed in the Bayesian learning cycle (Figure 1), a Bayesian reasoning framework allows scientists to integrate the strengths of both, inductive and deductive research approaches, in a formalized way. Further, when specifying the prior, scientists openly and explicitly need to describe the (un)certainty in the current state of knowledge and the posterior captures how the data has changed the knowledge and the (un)certainty in that knowledge. Thus, Bayesian reasoning is, at least in principle, open and transparent by design.

Figure 1
The Bayesian learning cycle adapted from Wagenmakers and Gronau (2018).
\includegraphics[scale = .75]{learning.png}

Having introduced Bayes theorem and an application of Bayes theorem as a cycle through which individuals can learn, we describe the role of uncertainty in science, with a focus on the role of probability.

\section{Uncertainty in Science and the Role of Probability}

Probability and uncertainty are ubiquitous in science—in scientific methodologies, the scientific concepts, or in how science is communicated (Gougis et al., 2016). The practice of science often begins with measurement, and with measurement comes measurement error, which has a random component. This random component arises from random variations in the measurement process and limits the certainty that one can have in the measurements. Reducing that measurement error has been critical for many discoveries in science, such as in sorting out the periodic table as we know it today (Fontani, 2014). However, deciding on adequate statistical procedures to determine measurement uncertainty regularly sparks debate, e.g., the recent study on the relationship between SARS-CoV-2 viral load and patient age by Jones et al. (2020) was heavily debated in the scientific community (Frick, 2020).

Probability and uncertainty are not only key concepts in science practices such as analyzing and interpreting data, but also part of the scientific concepts themselves. One example is Heisenberg’s uncertainty principle or quantum mechanics more generally  (Feynman, 1951). In quantum mechanics, confidence in measurement is limited not only by measurement error but in principle—even in the perfect experimental setup there is no deterministic outcome. Instead, one has to calculate the probability of each possible outcome using the Born rule. To this day, there is an ongoing argument among physicists and philosophers about how to interpret probability and the puzzling quantum mechanical phenomena such as quantum tunnelling (e.g., Caroll [2020]). Another example of how probability and uncertainty are central to learning scientific concepts is the probabilistic nature of evolution (Fiedler et al., 2017). Evolution is a probabilistic process as the emergence of new variations (e.g., through mutations) can only be described probabilistically. Further, the continued survival of these variations is governed by phenomena such as natural selection or random drift which also escape a deterministic description and are thus modelled probabilistically. Similar to within quantum physics, how to interpret this probability has also sparked debate among researchers (Millstein, 2014).

Given that probability and uncertainty spark debate among scientists, it is perhaps unsurprising that misinterpretations happen when scientists in the natural and engineering sciences communicate about their findings with their peers or the public (e.g., Gigerenzer 2004, McShane & Gal, 2017, Cumming, 2014). For example, error bars in the form of confidence intervals are routinely interpreted as distributions that assign higher probability to the center of the interval (Kruschke & Liddell, 2018a) or p-values are interpreted as effect sizes (Gelman & Carlin, 2017; Nuzzo; 2014) although p-values describe how (in)compatible a set of data is with a set of assumptions. In sum, misinterpretations of statements about probability and uncertainty are common across the range of modes in which scientists communicate about their work.

\section{Uncertainty in Science Education}

Given the above-noted roles of probability in scientific concepts and theories, it is unsurprising that educational researchers have documented how students learn about probability. As one example, probability and chance events play an important role in the life sciences (Garfield, 2003; Garvin-Doxas & Klymkowski, 2017), particularly in evolutionary processes and learning about them (Tibell & Harms, 2017). Recently, Fiedler et al. (2019) investigated the relation between statistical reasoning and acceptance and knowledge about evolution in a large sample of nearly 500 university students in the US. They found that students’ statistical reasoning capabilities are a strong predictor of both acceptance of evolution and knowledge about evolution. Thus, this study suggests (but does not directly show) that instructional strategies for supporting students’ statistical reasoning could improve their learning about specific evolutariony concepts.

In the physical sciences, students’ struggles with the probabilistic nature of quantum physics (Bao & Redish, 2002) and nuclear decay (Santostasi et al., 2017) are known, but the underlying reasons and mechanisms remain less researched than in evolution. Research into students’ misconceptions inspired by conceptual change research  (e.g., Marshman & Singh, 2015, Aguilar et al., 2014, Stefani & Tsarparlis, 2009) has started to address this gap. Findings seem to mirror the position of Batanero et al. (1994) in that the concepts of probability are taught too abstractly and in a way that does not build upon and connect with the ideas that students’ already hold, the experiences that students make, and the language that students use. Thus, these studies affirm students' struggles with scientific concepts that are probabilistic but hint at the underlying issue being foundational, conceptual struggles with ideas about probability not limited to science learning. Indeed, the language and representations used to communicate probability and uncertainty independent of specific scientific content domains or scientific practices are also known to present difficulties to learners. The work of Tversky and Kahneman (1974) points to a whole range of biases that the language in which statements about probability and uncertainty are framed can elicit. A prominent example is the neglect of base rate information in the (timely) context of medical testing, that is, the probability of having a medical condition on the basis of a positive test result is often hugely overestimated because information about the base rate of people affected from the condition is neglected (Kahneman, 2012).

More recent work by Gigerenzer and colleagues has provided important insights into the underlying mechanisms for the occurence of these biases (e.g., Gigerenzer & Hoffrange, 1995; Jenny et al., 2018). Their research suggests that the biases are at least partly the result of information about probability and uncertainty being provided in a way that is incompatible with the heuristics that people develop from their everyday experiences. Based on this assumption, Jenny et al. (2018) were able to demonstrate that students judge probability significantly better, when language is adopted that better aligns with the heuristics that most learners develop from their everyday experiences. In sum, changing representations of probability and uncertainty so that they better align with ideas and heuristics that learners hold so that these support instead of inhibit learning seems feasible.

\section{How Learners and People Make Sense of Data in Light of Uncertainty}

Recent work in cognitive and psychological science on how children learn argues that children understand and learn about the world through the intuitive application of Bayesian ideas (Gopnik, 2012; Tenenbaum et al., 2006; Griffiths et al., 2008). This idea has led current research to conceptualize human learning in terms of Bayesian probabilistic models of learning. In other words, people (and children) may view (or come to view) the world probabilistically, and in ways that happen to align with Bayesian interpretations of probability.

This view of how children learn from data has become especially prominent in developmental science and is successfully applied there (Gopnik and Tenenbaum, 2007; Gopnik & Wellman, 2012). One might assume that there has been a ready application of these ideas about how children think about the world in a Bayesian way to education. Developmentalists have pointed to the potential and opportunity here. Gopnik (2012) wrote that the use of Bayesian methods to understand child development can serve as “a scientific foundation for a long tradition on ‘inquiry-based’ science education,” that “could lead us to much more specific and scientifically supported proposals for education” (p. 1627).

Despite arguments for integrating developmental and educational research, it has largely not been the case that—with respect to advances in Bayesian methods—“science itself could help turn young children’s natural curiosity and brilliance into better science teaching and learning” (Gopnik, 2012, p. 1627), as some science education scholars have bemoaned (Lehrer & Schauble, 2006). Indeed, scholars have argued that it is often obsolete applications of Piaget’s ideas (Piaget & Inhelder, 1969) that represent the greatest contribution of developmental ideas to education (Lehrer & Schauble, 2007; Metz, 1995). We argue that this is still largely the case in the present, despite advances in both developmental research as well as statistical methodology related to Bayesian ideas—and that it is time to start exploring how bringing out the inherent Bayesian ways of thinking in education may provide support for learners’ efforts to make sense of the world.

Related science education research emphasizes the role of prior knowledge that students bring with them to the analysis of data and how that prior knowledge impacts the inferences they make about data. For example, Masnick et al. (2007) summarized findings across a number of studies showing that when elementary grade students reasoned with data in a context about which they had strong and valid prior understandings that informed their interpretation of data (about force and motion in the context of ramps), and they were able to explain sources of error in the data (Masnick & Klahr, 2003), but when the data was in a context they did not understand well, they struggled to distinguish between random variation due to sources of error and sources of variation due to something that was causing a difference in the data (generated by measuring the motion of pendulums). Later design- and laboratory-based research has shown that students can develop an understanding of scientific ideas through the analysis of data (Lehrer & Schauble, 2004; Masnick et al., 2017).

Laboratory-based research, especially, has examined how learners’ prior understanding may have a bearing on how they make sense of how likely different outcomes are (Klahr & Dunbar, 1998; Masnick et al., 2007; Schwartz et al., 2007). While these studies focused on how students’ prior ideas and data interact, Bayesian frameworks were not used to formalize this process of integrating data in light of prior beliefs about scientific phenomena. In short, this work aligns closely with the Bayesian models of how children learn from data described by Gopnik (2012), but this research does not explicitly interpret their findings in light of these ideas about how children learn from data in a Bayesian way.

Overall, developmental and psychological research into Bayesian models of cognition supports and bolsters efforts in science education to design instruction based around eliciting and understanding students’ ideas (Gotwals & Birmingham, 2012; Haverly et al., 2019; Windschitl et al., 2012). There is an opportunity to use Bayesian ideas as a framework for formalizing these ideas about how prior ideas affect students’ reasoning about data and how teachers can leverage students’ ideas to design instruction, yet Bayes has mostly been absent from research in science education.
After having reviewed how people learn about probability and uncertainty and where they face difficulties, we want to present how Bayes could be part of a solution. First, we will make the case how Bayes could support science literacy at large before delving into more concrete topics.

\section{Bayes and Science Literacy}

\subsection{Overview of Bayes and Science Literacy}

Having described the role of probability and uncertainty in science and science education, we now describe what these Bayesian ideas can mean for science literacy. We see two primary ways in which Bayesian methods can enhance science literacy (National Academies of Sciences, Engineering, and Medicine, 2016), or a range of capabilities that people may possess which have value (individually or collectively) to them. These capabilities range from foundational literacies, including numeracy and textual literacy (what is most thought of concerning literacy in common parlance), content knowledge, and epistemic knowledge and other ways of thinking and habits of mind (National Academies of Sciences, Engineering, and Medicine, 2016). One way that Bayesian ideas can inform discussions of science literacy is to provide individuals with a technique for combining their prior understanding with new evidence. Along this line, Bayes and the Bayesian learning cycle more specifically can provide a conceptual tool for people to combine their understanding with new evidence (and data) in a rigorous and coherent way, as we discuss next.

The history of science is, in many ways, a story of established ideas being built upon or supplanted (Fara, 2010), and the contemporary emphasis on viewing science as a practice was sparked by Kuhn’s (1962) historical-epistemological account on what this suggests about the nature and philosophy of science (Godfrey-Smith, 2009). Yet, our ways of interpreting scientific evidence do rarely recognize this, perhaps for good reason: Most of the time, new data or information do not contradict foundational scientific ideas (Kuhn, 1962). The same phenomenon, though, is present on a more micro-level: While scientists may be familiar with the data they collect not eliding with ideas that are a part of the scientific consensus, in media and the public discourse scientific ideas are often presented as fixed and certain—even in cases for which consensus has not been reached (Collins, 1987).

These historical and philosophical ideas align with a Bayesian perspective on science literacy. Established scientific ideas can be considered to be those that individual people - psychologically - should be more confident about, whereas newer ideas, or data suggesting new ideas, can be weighted less. In this way, Bayes can provide a simple heuristic for people to evaluate new information or evidence; by asking the question of how informative is the evidence in light of the prior. Thus, individuals possess a tool, one that is non-dichotomous, in that the evidence is not true or false, and also probabilistic, in that what individuals can think, going forward, is about the extent to which a particular belief has merit and is informative.

A concrete example can help to illustrate how a Bayesian perspective on a topic relevant to science literacy may offer a different way forward. Consider a salient topic, medical vaccines for diseases, such as Polio and the disease caused by the coronavirus SARS-CoV-2, COVID-19. Most evidence suggests that vaccines have been safe, historically, and are generally safe, at present (Centers for Disease Control and Prevention, 2021). Nevertheless, they are not perfectly safe, as unlikely, adverse outcomes do happen, and it is possible that there are adverse side effects from some vaccines that research has not yet revealed. In light of this evidence, consider the development of a new vaccine, such as those available at the time of this writing for COVID-19. Some use long-established methods of vaccine development (namely, deactivating SARS-CoV-2 particles but retaining cell signalling proteins to which humans develop an immune response—and, after, a degree of immunity) and injecting these deactivated virus particles into people), whereas others use a new mechanism (namely, injecting messenger RNA into people; this messenger RNA is then transcribed by cells in people’s bodies to produce a particular protein common to SARS-CoV-2 particles, which human’s bodies than learn to recognize to develop immunity.

How should we consider evidence about the vaccines? Prior to any data from medical clinical trials, it would be reasonable to consider these vaccines to probably be safe in light of the safety of other vaccines (and on the basis of similarities in their mechanisms of action), but to have a strong degree of caution about how safe they are. After initial trials were completed, the degree of safety that individuals associate with these vaccines—in light of the prior and the limited evidence—could increase, but still not be near the level of safety that many would consider acceptable. Following the completion of clinical trials that show minimal side effects (holding aside their efficacy), one’s estimates of the safety of these vaccines could increase by a substantial amount, but concerns could still remain, particularly for messenger RNA vaccines, which are newer, and may, plausibly, have unintended side effects.

\subsection{Strategies for Supporting Science Literacy}

What is the result of this entire process of weighting new evidence in light of a generic prior that we discussed above? We argue that there are three ways that Bayesian methods could support science literacy. 
First, the result is a non-binary inference, one that doesn’t speak to whether or not the vaccine is safe, but to what degree it is safe, doing so with a language of probability and uncertainty. Though a small difference, discussing the safety of vaccines on these terms can provide a more solid foundation for principled disagreements and arguments.

Second, Bayes provides a heuristic and a language for engaging with uncertainty in a science literacy context in a principled way that is in-line with how most of us likely already interpret data. Specifically Bayes provides people with a way to talk about weighing between ideas and their prior beliefs and new information and data. In short, Bayes gives people a chance to deal with uncertainty in a principled but accessible way.

Last, his process could be formalized mathematically.

I think there are several very general ideas that could be illustrated with Bayes’ rule. These include:

\begin{enumerate}
    \item When we gain new information, our updated knowledge of the world is a compromise between our old knowledge and the new information.
    \item Today’s posterior is tomorrow’s prior 
    \item Hypotheses/claims that predict new data well receive a boost in credibility, whereas those that predict the data poorly suffer a decline.
    \item All knowledge is relative
    \item Avoid absolute beliefs (Cromwell’s rule; these are not amenable to learning)
    \item Surprise lost is credibility gained
    \item Polya’s fundamental inductive pattern: “This inductive pattern says nothing surprising. On the contrary, it expresses a belief which no reasonable person seems to doubt: The verification of a consequence renders a conjecture more credible. With a little attention, we can observe countless reasonings in everyday life, in the law courts, in science, etc., which appear to confirm to our pattern.” (Polya, 1954, pp. 4-5)
    \item Extraordinary claims require extraordinary evidence
\end{enumerate}

\section{Bayes and Science Education}

\subsection{Overview of Bayes and Science Education}

Research on the role of Bayes in science education has mostly been carried out in the context of undergraduate statistics education, in which there have been a number of calls to action (Gould et al., 2018; Hoegh et al., 2020), examples and design options (Albert, 2002; Bolstad, 2002; Hu, 2020; Gelman, 2008; Witmer, 2017) and debate (Johnson et al., 2020) over how to advance the place of Bayesian methods in undergraduate statistics and data science degree programs. The accessibility of Bayesian methods in undergraduate classes is the result of advances in the necessary (for many uses) computer power (Gould et al., 2018) as well as the availability of tools that facilitate Bayesian analysis, especially for newcomers (Albert & Hu, 2020). As evidenced by the recent special issue of the Journal of Statistics Education of which several of the above-referenced articles are part, the pedagogy of Bayesian statistics is an active area of research in statistics education at the undergraduate level.

There is little research outside of statistics education in the broader science education community. Two publications merit mention though. The work of Szu and Osborne (2012) and Nussbaum (2011) both applied Bayesian perspectives to the science practice of argumentation. Szu and Osborne presented the case for how and why Bayes Theorem can apply to research and practice on scientific reasoning, broadly, and argumentation, particularly. This paper represents the most comprehensive account of the relevance of Bayesian methods for science education. Szu and Osborne argue that Bayes is useful as both a formal mathematical tool, and a conceptual one; indeed, they write that considering the degrees of certainty in beliefs that individual students hold—different from applications of Bayes Theorem that follow more or less deterministically from “external, objectively probabilistic systems” (p. 61)—is “the key leap that characterizes the debate about the value of Bayesian inference as a model of scientific reasoning” (p. 61). In this way, Szu and Osborne argue that the greatest use of Bayes theorem in science classrooms is as a model of informal scientific reasoning, aligning with similar (informal) approaches to inference within the statistics education research community (e.g., Batanero et al., 2016; Makar & Rubin, 2018).  They offer some research-related backing for the use of Bayes Theorem (e.g., noting how conceptual change research, particularly, and constructivist research, broadly, can both be explained coherently within a Bayesian approach) as well as some practical, instructional recommendations, some of which we detail later in this section.

Nussbaum’s (2011) work on Bayesian approaches to argumentation complement the work of Szu and Osborne in that Nussbaum describes both an application of Bayesian methods in K-12 classroom contexts and ideas about how Bayesian methods can serve as an analytic framework for students’ argumentation. Concerning the latter, Nussbaum describes how the social issue of raising taxes to provide resources to homeless individuals was a rich context for students to engage in forms of Bayesian reasoning. In this application, Nussbaum describes how the prior and likelihood could be obtained from empirical evidence—in this way, illustrating what Szu and Osborne (2012) characterize as the more externally objective use of Bayes Theorem—and how the estimates that result from applying Bayes theorem led students to re-evaluate their initial arguments.

In summary, the work of Szu and Osborne (2012) and Nussbaum begin to apply Bayesian ideas to science education and other relevant classroom contexts, making Bayesian ideas more vivid in the process. Curiously, no work has applied Bayesian methods to another science practice, that of analyzing and interpreting data; one for which Bayesian methods have clear conceptual applicability and utility, but one for which the absence of suitable tools may provide to be a barrier. In this way, Bayesian methods could serve as a cross-cutting concept (National Research Council, 2012) that unites the practice of argumentation with analyzing and interpreting data, and using mathematics and computational thinking.

\subsection{Strategies for Supporting Science Education}

Bayesian methods clearly have some utility in science education, even if that utility is under-realized at present. Next, we present some of the specific ways Bayesian methods could enhance science education—and could do so in a way that makes Bayes more tractable.

First, we should consider “unplugged” versions of Bayesian analyses, akin to how computer science education scholars have emphasized that computational thinking does not necessarily require the use of a computer (Yadav et al., 2018). These could interface with the science and engineering practice of modeling; could students use diagrammatic models in conjunction with Bayesian “heuristics” to revise their models in a similar manner to how this practice can unfold in classrooms (see Schwarz et al., 2009).

Another advance should be to build out curricular strategies, elaborating on the ideas of Szu and Osborne (2011) and Nussbaum (2011). At a high-level, Szu and Osborne (2011) write that in order to reason about data in light of their initial beliefs (and to construct a kind of informal likelihood statistic), students “need to see judgments about data and evidence being an assessment not only of the probability of the hypothesis being correct but also of it being wrong” (p. 67). Filling in more details, speaking from both the perspective of a researcher (of argumentation) and an educator (or, a designer of classroom activities), Nussbaum summarizes that using Bayes “forces one to (a) think about an issue from different perspectives, (b) reflect on one’s range of certainty and uncertainty, and (c) conduct a simple com- puter simulation (e.g., calculating the product) that integrates various facets of the issue together and affords further reflection” (p. 99).

These are helpful suggestions, and they complement finer-grained instructional techniques, especially those around how students can reason about uncertainty probabilistically. Particularly, Binder et al. (2018), for instance, suggest replacing probabilities with natural frequencies (in alignment with a great deal of work on how humans understand probability; Gigerenzer & Hoffrange, 1995) as well as using and interpreting tree diagrams to depict natural frequencies. Martignon and Erickson (2014)—building on much earlier experimental work by Wason and Shapiro (1966)—provide a simple example of how even fourth graders, with the right tools to represent probabilities, can support students to transition from logical to probabilistic reasoning. These could also support the aim of developing unplugged versions of Bayesian analyses.

Third, we should, of course, consider teachers and how teachers can (or already) support students to reason about uncertainty, and, possibly, to use Bayesian ideas. While some scholars have contributed foundational work demonstrating how this uncertainty can be productive as a part of teaching practice (Manz & Suarez, 2018), a broad and formal perspective on how uncertainty can be approached in science education is needed. There is the need to explore the potential of a principled approach to uncertainty, one that may be intuitive to learners and useful to teachers, but is as yet under-represented in the science education literature. If learners and people better understand uncertainty, they can better understand the scientific process which helps to build trust in science. Further, a better understanding of uncertainty can help to learn about scientific--and other--concepts where probability plays a key role; it may be that learning about how Bayes rule can be used in legal or medical contexts could impress upon students the cross-cutting nature of probability and uncertainty. This is especially helpful in light of the teacher's practice of noticing and responding to students' ideas and the process through which students change their understanding.

Finally, recently developed tools, particularly the JASP statistical software, may be helpful. JASP allows for both conventional frequentist as well as Bayesian versions of commonly used statistical tests, such as t-tests and regression analyses. There are a number of accessible guides available for specifying Bayesian versions of analyses (van Doorn et al., 2020; Wagenmakers et al., 2018), and there is a great potential in building particular examples within JASP (that are aligned to science education curricular standards) that could support science educators to have greater access to the power of Bayesian methods in their classrooms. Another possibly useful tool could be Bayesian networks (see Bayes Box; Bayes Box, 2021), which could be used to model complex system in a way that is similar to other tools used in science education for modeling complex systems, such as SageModeler (2020).

\section{Summary and Conclusion}

While some educational researchers may state that such theoretical, abstract ideas about the philosophical and statistical underpinnings of working with data are irrelevant to how students work with data, we make the case, here, that these ideas can inform instructional practice, assessments, and the design of activities that involve students in working with data, especially given some of the clear problems that have been articulated around the present frequentist Null Hypothesis Statistical Testing (NHST)-based approaches (Cumming, 2014; Gigerenzer, 2004).

Science education researchers may note that these ideas are conceptually-focused, rather than aligning with the present emphasis in science education upon engaging students in science and engineering practice in order to make sense of core ideas and to engage with cross-cutting concepts (National Research Council, 2012). We argue that Bayes is unique in that it, perhaps akin to a cross-cutting concept in the sense of a style of reasoning (Szu & Osborne, 2012; Osborne et al. 2018) that blends conceptual ideas with practices. Bayes is both a philosophical and psychological perspective on how to learn from data as well as a statistical framework. Indeed, a key potential benefit of Bayes may be that it provides a framework for helping students to transition from their conceptual understanding of scientific ideas to quantitatively analyzing data.

In this paper, we have introduced Bayes as a powerful way of conceptualizing probability and uncertainty. These are two concepts that are ubiquitous in the learning, doining, and communicating of science and therefore critical aspects of science literacy (National Academies of Sciences, Engineering, and Medicine, 2016). However, probability and uncertainty are also infamously challenging concepts—for experts and laypersons alike (Gigerenzer et al. 2004; Sedlmeier, 2007; Tversky & Kahneman, 1977). Thus, it is not surprising when the public misinterprets statements about probability and uncertainty but the consequences may be dire when trust in science is eroded, and, sadly, examples of this are plentiful during current the COVID-19, e.g., the efficiency of masks in preventing the spread of COVID and the efficiency of the AstraZeneca vaccine were strongly questioned, delaying efforts to comb at the pandemic. The public, however, can hardly be blamed for this in light of the prevalent findings that the frequentist conceptualizations of probability and uncertainty in school are challenging and uncommonly align with or build on peoples’ intuitions. As research from cognitive and psychological science as well as emerging work from statistics and statistics data science education demonstrates, conceptualizing probability and uncertainty from a Bayesian perspective better aligns with peoples’ intuitions (Kruschke & Liddell, 2018a; Szu & Osborne, 2012) about these concepts and thus has potential to support learning about probability and uncertainty in science.

We have outlined numerous vantage points for research agendas to explore the potential of Bayes in science education so that future research can identify how students can become better reasoners in light of uncertainty. We have also discussed how we as scientists and science educators can contribute to the public trust in science by becoming better in how we communicate about probability and uncertainty. In all, Bayes is a big idea—one with epistemological and philosophical, statistical, and pedagogical implications. We argue that the time for advancing Bayesian methods has come, particularly in light of distrust in science and science education and the way Bayesian methods can address this state of affairs by offering a different foundation and language around a core but sometimes unsettling tenet and persistent feature, uncertainty.

\printbibliography

\appendix

\section{code}
add JASP links/code/etc.?

\end{document}

%% 
%% Copyright (C) 2019 by Daniel A. Weiss <daniel.weiss.led at gmail.com>
%% 
%% This work may be distributed and/or modified under the
%% conditions of the LaTeX Project Public License (LPPL), either
%% version 1.3c of this license or (at your option) any later
%% version.  The latest version of this license is in the file:
%% 
%% http://www.latex-project.org/lppl.txt
%% 
%% Users may freely modify these files without permission, as long as the
%% copyright line and this statement are maintained intact.
%% 
%% This work is not endorsed by, affiliated with, or probably even known
%% by, the American Psychological Association.
%% 
%% This work is "maintained" (as per LPPL maintenance status) by
%% Daniel A. Weiss.
%% 
%% This work consists of the file  apa7.dtx
%% and the derived files           apa7.ins,
%%                                 apa7.cls,
%%                                 apa7.pdf,
%%                                 README,
%%                                 APA7american.txt,
%%                                 APA7british.txt,
%%                                 APA7dutch.txt,
%%                                 APA7english.txt,
%%                                 APA7german.txt,
%%                                 APA7ngerman.txt,
%%                                 APA7greek.txt,
%%                                 APA7czech.txt,
%%                                 APA7turkish.txt,
%%                                 APA7endfloat.cfg,
%%                                 Figure1.pdf,
%%                                 shortsample.tex,
%%                                 longsample.tex, and
%%                                 bibliography.bib.
%% 
%%
%% End of file `./samples/longsample.tex'.